Left-linear rules in adhesive categories

### 2024-02-08 icalp/fusione.tex
general pass

@@

\item Three-steps lemma (Lemma~\ref{lem:iig1}), which is a weakened version of ``indepedence is global''.
  \begin{itemize}

@@

\rem{Mi pare si possa omettere}{
  If we look to direct derivations as
  transitions, it is natural to consider them as edges in a direct
  graph. Taking objects as vertices objects led us to the following
  definition \cite{heindel2009category}.

  \begin{definition}
    Let $(\X, \R)$ be a left-linear DPO rewrityng system, with $\X$
    $\mathcal{M}$-adhesive. The \emph{DPO-derivation graph} of
    $(\X, \R)$ is the (large) directed graph $\gpo$ having as vertices
    the objects of $\X$ and in which an edge between $G$ and $H$ is a
    direct derivation $\dder{D}\colon G\Mapsto H$. A \emph{derivation}
    $\der{D}$ between two objects $G$ and $H$ is a path between them
    in $\gpo$. The \emph{source} and \emph{target} of $\der{D}$ are,
    respectively, $G$ and $H$.
  \end{definition}
}
\begin{definition}[Derivation]
  Let $(\X, \R)$ be a left-linear DPO rewrityng system, with $\X$
  $\mathcal{M}$-adhesive.  An \emph{empty derivation} starting and
  ending in $G$ is just $G$ itself.  A \emph{non-empty derivation}
  $\dder{D}$ is a sequence $\{\dder{D}_i\}_{i=0}^n$ of direct
  derivations such that:
  \begin{enumerate}
  \item for every index $i$, $\dder{D}_i$ is a direct derivation $G_i \Mapsto G_{i+1}$;
  \item $G_0=G$ and $G_{n+1}=H$.
  \end{enumerate}

  We will call the number $n+1$ the \emph{length} of the derivation,
  denoted by $\lgh(\der{D})$. We will also say that an empty
  derivation has length $0$.

  Moreover, if every $\dder{D}_i$ applies the rule $\rho_i\in R$, then
  we can define an associated sequence of rules as $r(\der{D})$ as
  $\{\rho_i\}_{i=0}^n$.
\end{definition}

\begin{remark}
  \label{rem:func}
  Consider a derivation $\der{D}$ in a left-linear DPO rewriting
  system $(\X, \R)$. We can take the subcategory $\Delta(\der{D})$ of
  $\X$ given by the arrows appearing in $\der{D}$. This subcategory
  comes equipped with an inclusion functor
  $I(\der{D})\colon \Delta(\der{D})\to \X$. Moreover, we can further
  define $\Deltamin(\der{D})$ as the subcategory of $\Delta(\der{D})$
  containing only the bottom row of the derivation.
\end{remark}

\noindent
\parbox{10cm}{%
  \begin{notation}
    Let $\der{D}=\{\dder{D}_i\}_{i=0}^n$ be a derivation. We will
    depict the $i^\text{th}$ element $\dder{D}_i$ of $\der{D}$ as in
    the diagram on the right. Notice that, in particular, if $\der{D}$
    is a derivation between $G$ and $H$, then $G_0=G$ and
    $G_{n+1}=H$. When $\der{D}$ has length $1$ we will suppress the
    indexes. In such case, we will also identify $\der{D}$ with its
    only element.
  \end{notation} }
\parbox{3cm}{
  $\xymatrix{L_i \ar[d]_{m_i}& K_i
    \ar[d]^{k_i}\ar@{>->}[l]_{l_i} \ar[r]^{r_i} & R_i \ar[d]^{h_i}\\
    G_{i} & \ar@{>->}[l]^{f_{i}} D_{i} \ar[r]_{g_{i}}& G_{i+1} }$
}

@@

\section{Switch equivalence }\label{sec:equi}

\todo{A VERY NICE INTRO}

@@

\begin{definition}[Switch]
  \label{def:switch}
  Let $(\X, \R)$ be a left-linear DPO rewriting system with $\X$ an
  $\mathcal{M}$-adhesive category and suppose that
  $\der{D}=\{\dder{D}_i\}_{i=0}^1$ is a derivation of length $2$ made
  by two weakly sequentially independent derivations $\dder{D}_0$,
  $\dder{D}_1$, with independence pair $(i_0, i_1)$:
  \[\xymatrix@C=15pt{L_0 \ar[d]_{m_0}&& K_0
      \ar[d]_{k_0}\ar@{>->}[ll]_{l_0} \ar[r]^{r_0} & R_0
      \ar@/^.35cm/[drrr]|(.3)\hole_(.4){i_0} \ar[dr]|(.3)\hole_{h_0}
      && L_1 \ar@/_.35cm/[dlll]^(.4){i_1} \ar[dl]|(.3)\hole^{m_1}& K_1
      \ar[d]^{k_1}\ar@{>->}[l]_{l_1} \ar[rr]^{r_1} && R_1
      \ar[d]^{h_1}\\G_0 && \ar@{>->}[ll]^{f_0} D_0 \ar[rr]_{g_0}&& G_1
      && \ar@{>->}[ll]^{f_1} D_1 \ar[rr]_{g_1}&& G_2}\]

  Take another derivation $\der{E}=\{\dder{E}_i\}_{i=0}^1$ as below
  \[\xymatrix@C=18pt{L_{\der{E},0} \ar[d]_{m_{\der{E},0}}&&
      K_{\der{E},0}
      \ar[d]_{k_{\der{E},0}}\ar@{>->}[ll]_{l_{\der{E},0}}
      \ar[r]^{r_{\der{E},0}} & R_{\der{E},0} \ar[dr]_{h_{\der{E},0}}
      && L_{\der{E},1} \ar[dl]^{m_{\der{E},1}}& K_{\der{E},1}
      \ar[d]^{k_{\der{E},1}}\ar@{>->}[l]_{l_{\der{E},1}}
      \ar[rr]^{r_{\der{E},1}} && R_{\der{E},1}
      \ar[d]^{h_{\der{E},1}}\\G_{0} && \ar@{>->}[ll]^{f_{\der{E},0}}
      D_{\der{E},0} \ar[rr]_{g_{\der{E},0}}&& G_{\der{E},1} &&
      \ar@{>->}[ll]^{f_{\der{E},1}} D_{\der{E},1}
      \ar[rr]_{g_{\der{E},1}}&& G_{2}}\] We say that $\der{E}$ is a
  \emph{switch} of $\dder{D}_0$ and $\dder{D}_1$ along $(i_1, i_2)$ if
  \begin{enumerate}
  \item $\dder{E}_0$ applies the rule used by $\dder{D}_1$ and
    $\dder{E}_1$ the one used by $\dder{D}_0$, so that
    \[(l_{\der{E},0}, r_{\der{E},0})=(l_1,r_1) \qquad (l_{\der{E},1},
      r_{\der{E},1})=(l_0,r_0)\]
  \item there exists an independence pair $(j_0, j_1)$ between
    $\dder{E}_0$ and $\dder{E}_1$ such that
    \[ m_0=f_{\der{E},0} \circ j_1 \qquad h_1=g_{\der{E},1} \circ
      j_0\]
  \item $m_{\der{E},0}= f_0\circ i_1$ and
    $h_{\der{E},1}= g_{1}\circ i_0$.
  \end{enumerate}

  If a switch of $\dder{D}_0$ and $\dder{D}_1$ exists we will say that
  they are \emph{switchable}.
\end{definition}

The fact that sequential indepedence implies switchability always holds for linear rules. [REF]
The result is so fundamental that it is not even stated, in the sense that switchability is not introduced axiomatically as in defition~\ref{def:switch}, but it is based on the explicit construction of a switch.

@@

\rem{remove}{
  \begin{remark}
    \label{rem:fact}
    Notice that, since $f_{\der{E}, 0}$ is in $\mathcal{M}$, there is
    at most one $j_1\colon L_0\to D_{\der{E},0}$ such that
    $m_0=f_{\der{E}, 0} \circ j_1$. Moreover, we already know that
    there is at most one $j_0\colon R_1\to D_{\der{E},1}$ such that
    $m_{\der{E},1}=f_{\der{E},1}\circ j_0$, therefore, the
    independence pair $(j_0,j_1)$ between the component of a switch is
    uniquely determined.
  \end{remark}
}

@@

  Take now the set $S:=\{a,b,c\}$ and endow it with the smallest order
  such that $a\leq b$ and $a\leq c$.  Take also the set $\mathbb{N}$
  ordered by $\geq$, so that $0$ is its maximum.  Let $\X$ be the
  category associated to the order
  $(S, \leq)\oplus (\mathbb{N}, \geq)$ which, by \Cref{rem:iso} is
  $\mathsf{Iso}$-adhesive. Notice that in this category the arrows
  $(a,b)\colon a\to b$ $(a,c)\colon a\to c$ do not have a pushout
  because $b$ and $c$ do not have a supremum.  Define $\R$ as the set
  containing the following two rules
  \[\xymatrix{a & a \ar[r]^{(a, 0)} \ar[l]_{\id{a}} & 0 & a & a
      \ar[r]^{(a, b)} \ar[l]_{\id{a}} & b }\]
  We can then consider the following derivation
  $\der{D}=\{\dder{D}_i\}_{i=0}^1$ of length $2$, made by two
  sequential independent direct derivations.
  \[\xymatrix@C=30pt{a \ar[d]_{m_0}&& a \ar[d]_{k_0}\ar[ll]_{\id{a}}
      \ar[r]^{(a,0)} & 0 \ar@/^.35cm/[drrr]|(.315)\hole_(.45){\id{0}}
      \ar[dr]|(.3)\hole_{\id{0}} && a \ar@/_.35cm/[dlll]^(.45){(a,c)}
      \ar[dl]|(.3)\hole^{(a,0)}& a \ar[d]^{(a,0)}\ar@{>->}[l]_{\id{a}}
      \ar[rr]^{(a,b)} && b \ar[d]^{(b,0)}\\c &&
      \ar@{>->}[ll]^{(\id{c})} c \ar[rr]_{(c,0)}&& 0 &&
      \ar@{>->}[ll]^{\id{0}} 0 \ar[rr]_{\id{0}}&& 0}\]

	\noindent
	\parbox{10cm}{ \hspace{15pt}On the other hand, the rule
          applied by $\dder{D}_1$ cannot be applied to $c$. indeed the
          unique morphism $a\to c$ is $(a,c)$, thus we get a diagram
          aside. But $b$ and $c$ do not have a supremum in the poset
          underlying $\X$, thus we do not get a direct derivation from
          $c$.}  \parbox{3cm}{
          \vspace{-.3cm}$\xymatrix{a \ar[d]_{(a,c)}& a
            \ar[d]_{(a,c)}\ar@{>->}[l]_{\id{a}} \ar[r]^{(a,b)} & b \\c
            & c \ar[l]^{\id{c}} }$}

      \end{example}

\rem{tenere solo una frase e risultati in appendice}{
Consider a switch $\der{E}=\{\dder{E}_i\}_{i=0}^1$ for $\dder{D}_0$ and $\dder{D}_1$, then by definition $\dder{E}_0$ and $\dder{E}_1$ are weakly sequentially independent. We can then wonder if they are switchable.

\begin{theoremEnd}[category=sec3]{proposition}
  \label{prop:switch}
  Let $\der{D}=\{\dder{D}_i\}_{i=0}^1$ be a derivation and suppose
  that $\der{E}=\{\dder{E}_i\}_{i=0}^1$ is a switch for it. Then
  $\der{D}$ is a switch for $\dder{E}_0$ and $\dder{E}_1$ along
  $(j_0, j_1)$.
\end{theoremEnd}

\begin{proof}
  \begin{proofEnd}
    The first point of \Cref{def:switch} is obviously satisfied. The
    second one follows at once since
    \[m_{\der{E},0}= f_0\circ i_1 \qquad h_{\der{E},1}= g_{1}\circ i_0\]
    Finally, the third point holds because we already know that
    \[m_{\der{E},0}= f_0\circ i_1 \qquad h_{\der{E},1}= g_{1}\circ i_0\]
  \end{proofEnd}
\end{proof}

Switches are unique up to abstraction equivalence, as shown by the next proposition.

\begin{theoremEnd}[category=sec3]{lemma}[Uniqueness of switches]\label{thm:switch_uni} Let $\der{D}=\{\dder{D}_{i}\}_{i=0}^1$ be a derivation and suppose that both $\der{E}=\{\dder{E}_i\}_{i=0}^1$  and $\der{F}=\{\dder{F}_i\}_{i=0}^1$ are switches of $\dder{D}_0$ and $\dder{D}_1$. Then $\der{E}$ and $\der{F}$ are abstraction equivalent.
\end{theoremEnd}
\begin{proof}
  \begin{proofEnd}
    By definition of switch we know that $\der{E}_0$ and $\der{F}_0$ have the same match $f_0\circ i_1$ and that $\der{E}_1$ and $\der{F}_1$ have the same comatch $g_1\circ i_0$. Thus we can build the solid part of the diagram below.
    \[\xymatrix@C=22pt{G_{0} && \ar@{>->}[ll]_{f_{\der{F},0}} D_{\der{F},0} \ar[rr]^{g_{\der{F},0}}&& G_{\der{F},1} && \ar@{>->}[ll]_{f_{\der{F},1}} D_{\der{F},1} \ar[rr]^{g_{\der{F},1}}&& G_{2}\\L_1 \ar[d]^{f_0\circ i_1} \ar[u]_{f_0\circ i_1}&& K_{1} \ar[d]^{k_{\der{E},0}} \ar[u]_{k_{\der{F},0}}\ar@{>->}[ll]_{l_1} \ar[r]^{r_1} & R_{1} \ar@/^.35cm/[drrr]|(.3)\hole_(.4){j_0} \ar[dr]|(.3)\hole_{h_{\der{E},0}} \ar@/_.35cm/[urrr]|(.3)\hole^(.4){a_0} \ar[ur]|(.3)\hole^{h_{\der{F},0}}& &L_{0} \ar@/_.35cm/[dlll]^(.4){j_1} \ar[dl]|(.3)\hole^{m_{\der{E},1}} \ar@/^.35cm/[ulll]_(.4){a_1} \ar[ul]|(.3)\hole_{m_{\der{F},1}}& K_{0} \ar[d]_{k_{\der{E},1}} \ar[u]^{k_{\der{F},1}}\ar@{>->}[l]_{l_{0}} \ar[rr]^{r_0} && R_{0} \ar[d]_{g_1\circ i_0} \ar[u]^{g_1\circ i_0}\\ \ar@/^.5cm/[uu]^{\id{G_0}} G_{0} && \ar[ll]^{f_{\der{E},0}} D_{\der{E},0} \ar@{.>}@/^.5cm/[uu]^(.33){\phi_{0}}|\hole \ar[rr]_{g_{\der{E},0}}&& G_{\der{E},1} \ar@{.>}[uu]^(.76){\psi_1}|(.45)\hole |(.55)\hole&& \ar@{>->}[ll]^{f_{\der{E},1}} D_{\der{E},1} \ar@{.>}@/_.5cm/[uu]_(.33){\phi_{1}}|\hole\ar[rr]_{g_{\der{E},1}}&& G_{2}\ar@{.>}@/_.5cm/[uu]_{\psi_2}}\]
    Now, by \Cref{prop:unique} we get $\phi_0\colon D_{\der{E},0}\to D_{\der{F},0}$ and $\psi_1 \colon G_{\der{E},1}\to G_{\der{F},1}$ such that
    \[\xymatrix{G_{0} & \ar@{>->}[l]_{f_{\der{F},0}} D_{\der{F},0} \ar[r]^{g_{\der{F},0}}& G_{\der{F},1} \\L_1 \ar[d]^{f_0\circ i_1} \ar[u]_{f_0\circ i_1}& K_{1} \ar[d]^{k_{\der{E},0}} \ar[u]_{k_{\der{F},0}}\ar[l]_{l_1} \ar[r]^{r_1} & R_{1}  \ar[d]_{h_{\der{E},0}}  \ar[u]^{h_{\der{F},0}}\\ \ar@/^.5cm/[uu]^{\id{G_0}} G_{0} & \ar@{>->}[l]^{f_{\der{E},0}} D_{\der{E},0} \ar@/^.5cm/[uu]^(.33){\phi_{0}}|\hole \ar[r]_{g_{\der{E},0}}& G_{\der{E},1} \ar@/_.5cm/[uu]_{\psi_1}}\]
    commutes. But then we have
    \[
      f_{\der{F},0}\circ \phi_0\circ j_1=f_{\der{E},0} \circ j_1=m_0=f_{\der{F},0}\circ a_1\]
    which entails $a_1= \phi_0\circ j_1$. This, in turn, gives us that
    \[\psi\circ m_{\der{E},1}=\psi \circ g_{\der{E},0}\circ j_1=g_{\der{F},0}\circ \phi_0\circ j_1=g_{\der{F},0}\circ a_1=m_{\der{F},1}\]

    Now, the square
    \[\xymatrix@C=30pt{K_0 \ar[d]_{k_{\der{E,1}}} \ar@{>->}[r]^{l_0}& L_0 \ar[d]^{\psi_1\circ m_{\der{E},1}} \\D_{\der{E},1} \ar@{>->}[r]_{\psi_1\circ f_{\der{E},1}}& G_{\der{F}, 1}} \]
    is a pushout and $\psi_1\circ m_{\der{E},1}=m_{\der{F},1}$, thus, by \Cref{prop:unique} there are $\phi_1\colon D_{\der{E},1}\to D_{\der{F},1}$ and $\psi_2\colon G_2\to G_2$ fitting in the diagram below
    \[\xymatrix{G_{\der{F},1} & \ar@{>->}[l]_{f_{\der{F},1}} D_{\der{F},1} \ar[r]^{g_{\der{F},1}}& G_2 \\L_0 \ar[d]^{m_{\der{E},1}} \ar[u]_{m_{\der{F},1}}& K_{1} \ar[d]^{m_{\der{E},1}} \ar[u]_{m_{\der{F},1}}\ar[l]_{l_1} \ar[r]^{r_1} & R_{1}  \ar[d]_{g_1\circ i_0}  \ar[u]^{g_1\circ i_0}\\ \ar@/^.5cm/[uu]^{\psi_1} G_{\der{E},1} & \ar@{>->}[l]^{f_{\der{E},1}} D_{\der{E},1} \ar@/^.5cm/[uu]^(.33){\phi_{1}}|\hole \ar[r]_{g_{\der{E},1}}& G_2 \ar@/_.5cm/[uu]_{\psi_2}}\]
    To conclude, it is now enough to notice that
    \[f_{\der{F},1} \circ \phi_1\circ j_0=\psi_1\circ f_{\der{E},1}\circ j_0=\psi_1\circ h_{\der{E,0}}=h_{\der{F},0}=f_{\der{F},1}\circ a_0\]
    allowing us to deduce $ \phi_1\circ j_0=a_0$.
  \end{proofEnd} \qedhere
\end{proof}
}

\rem{appendice?}
{We can now derive another property  of switches.

>>>>

% full or conference version
\newif\iffull
\fulltrue
%\fullfalse

% additional text
\iffull
\newcommand{\full}[1]{#1}
%\newcommand{\short}[1]{}
\else 
\newcommand{\full}[1]{}
%\newcommand{\short}[1]{#1}
\fi

@@

\item Three-steps lemma (Lemma~\ref{lem:indep-global-left}), which is a weakened version of ``indepedence is global''.
  \begin{itemize}

@@

\full{
  %
  If we look to direct derivations as
  transitions, it is natural to consider them as edges in a direct
  graph. Taking objects as vertices objects led us to the following
  definition \cite{heindel2009category}.

  \begin{definition}
    Let $(\X, \R)$ be a left-linear DPO rewrityng system, with $\X$
    $\mathcal{M}$-adhesive. The \emph{DPO-derivation graph} of
    $(\X, \R)$ is the (large) directed graph $\gpo$ having as vertices
    the objects of $\X$ and in which an edge between $G$ and $H$ is a
    direct derivation $\dder{D}\colon G\Mapsto H$. A \emph{derivation}
    $\der{D}$ between two objects $G$ and $H$ is a path between them
    in $\gpo$. The \emph{source} and \emph{target} of $\der{D}$ are,
    respectively, $G$ and $H$.
  \end{definition}
}

A derivation can now defined simply as a sequence of direct derivations. More precisely:

\begin{definition}[Derivation]
  Let $(\X, \R)$ be a left-linear DPO rewrityng system, with $\X$
  $\mathcal{M}$-adhesive. Given $G, H \in \X$, a \emph{derivation}
  $\der{D}$ with source $G$ and target $H$, written
  $\der{D}: G \Mapsto H$ is defined as a sequence
  $\{\dder{D}_i\}_{i=0}^n$ of direct derivations such that
  \begin{enumerate}
  \item for every index $i$, $\dder{D}_i$ is a direct derivation $G_i \Mapsto G_{i+1}$;
  \item $G_0=G$ and $G_{n+1}=H$.
  \end{enumerate}
  Moreover, for each $G \in \X$ we have $G : G \Mapsto G$ is a
  derivation, called an \emph{empty derivation}.

  We will call the number $n+1$ the \emph{length} of the derivation,
  denoted by $\lgh(\der{D})$. We will also say that an empty
  derivation has length $0$.\todo{P: used}

  Moreover, if every $\dder{D}_i$ applies the rule $\rho_i\in R$, then
  we can define an associated sequence of rules as $r(\der{D})$ as
  $\{\rho_i\}_{i=0}^n$.\todo{P: Used?}
\end{definition}

Derivations naturally compose: given $\der{D}: G \Mapsto H$ and $\der{E}: H \Mapsto F$ we can consider their composition $\der{D} \cdot \der{E}: G \to F$, defined in the obvious way.

\todo[inline]{P: se non impatta sul resto forse sarebbe forse piu' carino definire:\\
  - for each $G \in \X$, $G : G \Mapsto G$ is a derivation\\  
  - for each direct derivation $\dder{D}: G \Mapsto H$ and derivation $\der{E} : H \Mapsto F$ we have that $\dder{D}; \der{E} : G \Mapsto F$ is a derivation
}

\rem{Suppongo questo serva quando poi si considerano le permutazioni consistenti, spostare? }{
\begin{remark}
  \label{rem:func}
  Consider a derivation $\der{D}$ in a left-linear DPO rewriting
  system $(\X, \R)$. We can take the subcategory $\Delta(\der{D})$ of
  $\X$ given by the arrows appearing in $\der{D}$. This subcategory
  comes equipped with an inclusion functor
  $I(\der{D})\colon \Delta(\der{D})\to \X$. Moreover, we can further
  define $\Deltamin(\der{D})$ as the subcategory of $\Delta(\der{D})$
  containing only the bottom row of the derivation.\todo{Used?}
\end{remark}

\noindent
\parbox{10cm}{%
  \begin{notation}
    Let $\der{D}=\{\dder{D}_i\}_{i=0}^n$ be a derivation. We will
    depict the $i^\text{th}$ element $\dder{D}_i$ of $\der{D}$ as in
    the diagram on the right. Notice that, in particular, if $\der{D}$
    is a derivation between $G$ and $H$, then $G_0=G$ and
    $G_{n+1}=H$. When $\der{D}$ has length $1$ we will suppress the
    indexes. In such case, we will also identify $\der{D}$ with its
    only element.
  \end{notation} }
\parbox{3cm}{
  $\xymatrix{L_i \ar[d]_{m_i}& K_i
    \ar[d]^{k_i}\ar@{>->}[l]_{l_i} \ar[r]^{r_i} & R_i \ar[d]^{h_i}\\
    G_{i} & \ar@{>->}[l]^{f_{i}} D_{i} \ar[r]_{g_{i}}& G_{i+1} }$
}
}

@@

\section{Independence in rewriting}
\label{sec:equi}

@@

\begin{definition}[Switch]
  \label{def:switch}
  Let $(\X, \R)$ be a left-linear DPO rewriting system with $\X$ an
  $\mathcal{M}$-adhesive category and suppose that
  $\der{D}=\{\dder{D}_i\}_{i=0}^1$ is a derivation of length $2$ made
  by two sequentially independent derivations $\dder{D}_0$,
  $\dder{D}_1$, with independence pair $(i_0, i_1)$, as in the diagram
  on the left below. A \emph{switch} of $\der{D}$ along $(i_0,i_1)$ is
  a derivation $\der{E}=\{\dder{E}_i\}_{i=0}^1$, using the same rules
  in reverse order, as in the diagram on the right below, 
  \[
    \def\objectstyle{\scriptstyle}\def\labelstyle{\scriptstyle}
    \xymatrix@C=1.1mm@R=8mm{
      {L_0} \ar[d]_{m_0}
      & & {K_0} \ar@{>->}[ll]_{l_0} \ar[rr]^{r_0} \ar[d]_{k_0}
      & & {R_0} \ar[drr]_{h_0}  \ar@(r,ul)@{.>}^(.75){i_0}[drrrrrr]
      & & & & 
      % 
      {L_1}\ar[dll]^{m_1} \ar@(l,ur)@{.>}_(.75){i_1}[dllllll]
      & & {K_1} \ar@{>->}[ll]_{l_1} \ar[rr]^{r_1} \ar[d]^{k_1}
      & & {R_1} \ar[d]^{h_1} \\
      % 
      % 
      {G}
      & & {D_0} \ar@{>->}[ll]^{f_0} \ar[rrrr]_{g_0}
      & & & & {H} & &
      & &  {D_1} \ar@{>->}[llll]^{f_1} \ar[rr]_{g_1}
      & & {H}
    }
    % 
    \xymatrix@C=1.1mm@R=8mm{
      {L_1} \ar[d]_{m_1'}
      & & {K_1} \ar@{>->}[ll]_{l_1} \ar[rr]^{r_1} \ar[d]_{k_1'}
      & & {R_1} \ar[drr]_{h_1'}  \ar@(r,ul)@{.>}^(.75){i_1'}[drrrrrr]
      & & & & 
      % 
      {L_0}\ar[dll]^{m_0'} \ar@(l,ur)@{.>}_(.75){i_0'}[dllllll]
      & & {K_0} \ar@{>->}[ll]_{l_0} \ar[rr]^{r_0} \ar[d]^{k_0'}
      & & {R_0} \ar[d]^{h_0'} \\
      % 
      % 
      {G}
      & & {D_1'} \ar@{>->}[ll]^{f_1'} \ar[rrrr]_{g_1'}
      & & & & {H'} & &
      & &  {D_0'} \ar@{>->}[llll]^{f_0'} \ar[rr]_{g_0'}
      & & {H'}  }
    %
  \]
  such that there is an independence pair $(i_1', i_0')$ between
  $\dder{E}_0$ and $\dder{E}_1$ and 
  \begin{center}   
    $m_0=f_0' \circ i_0'$
    \qquad $h_1=g_1' \circ i_1'$
    \qquad $m_0'= f_0 \circ i_1$
    \qquad $h_1'= g_{1}\circ i_0$.
  \end{center}

  If a switch of $\dder{D}_0$ and $\dder{D}_1$ exists we will say that
  they are \emph{switchable}.

  % \[
  %   \xymatrix@C=7pt{L_0 \ar[d]_{m_0}&& K_0
  %     \ar[d]_{k_0}\ar@{>->}[ll]_{l_0} \ar[r]^{r_0} & R_0
  %     \ar@/^.35cm/[drrr]|(.3)\hole_(.4){i_0} \ar[dr]|(.3)\hole_{h_0}
  %     && L_1 \ar@/_.35cm/[dlll]^(.4){i_1} \ar[dl]|(.3)\hole^{m_1}& K_1
  %     \ar[d]^{k_1}\ar@{>->}[l]_{l_1} \ar[rr]^{r_1} && R_1
  %     \ar[d]^{h_1}\\G_0 && \ar@{>->}[ll]^{f_0} D_0 \ar[rr]_{g_0}&& G_1
  %     && \ar@{>->}[ll]^{f_1} D_1 \ar[rr]_{g_1}&& G_2}\]

  % Take another derivation $\der{E}=\{\dder{E}_i\}_{i=0}^1$ as below
  % \[\xymatrix@C=18pt{L_{\der{E},0} \ar[d]_{m_{\der{E},0}}&&
  %     K_{\der{E},0}
  %     \ar[d]_{k_{\der{E},0}}\ar@{>->}[ll]_{l_{\der{E},0}}
  %     \ar[r]^{r_{\der{E},0}} & R_{\der{E},0} \ar[dr]_{h_{\der{E},0}}
  %     && L_{\der{E},1} \ar[dl]^{m_{\der{E},1}}& K_{\der{E},1}
  %     \ar[d]^{k_{\der{E},1}}\ar@{>->}[l]_{l_{\der{E},1}}
  %     \ar[rr]^{r_{\der{E},1}} && R_{\der{E},1}
  %     \ar[d]^{h_{\der{E},1}}\\G_{0} && \ar@{>->}[ll]^{f_{\der{E},0}}
  %     D_{\der{E},0} \ar[rr]_{g_{\der{E},0}}&& G_{\der{E},1} &&
  %     \ar@{>->}[ll]^{f_{\der{E},1}} D_{\der{E},1}
  %     \ar[rr]_{g_{\der{E},1}}&& G_{2}}\] We say that $\der{E}$ is a
  % \emph{switch} of $\dder{D}_0$ and $\dder{D}_1$ along $(i_1, i_2)$ if
  % \begin{enumerate}
  % \item $\dder{E}_0$ applies the rule used by $\dder{D}_1$ and
  %   $\dder{E}_1$ the one used by $\dder{D}_0$, so that
  %   \[(l_{\der{E},0}, r_{\der{E},0})=(l_1,r_1) \qquad (l_{\der{E},1},
  %     r_{\der{E},1})=(l_0,r_0)\]
  % \item there exists an independence pair $(j_0, j_1)$ between
  %   $\dder{E}_0$ and $\dder{E}_1$ such that
  %   \[ m_0=f_{\der{E},0} \circ j_1 \qquad h_1=g_{\der{E},1} \circ
  %     j_0\]
  % \item $m_{\der{E},0}= f_0\circ i_1$ and
  %   $h_{\der{E},1}= g_{1}\circ i_0$.
  % \end{enumerate}

  % If a switch of $\dder{D}_0$ and $\dder{D}_1$ exists we will say that
  % they are \emph{switchable}.
\end{definition}

As already mentioned, when viewing derivations as concurrent computation, for linear graph rewriting systems it is canonical to identify  derivations which are equal ``up to switching''. The same notion can be given for left-linear rewriting systems relying on notion of switch.

\begin{definition}[switch equivalence]
  \label{de:switch-equivalence}
  Let $(\X, \R)$ be a left-linear DPO rewriting system.  Let
  $\der{D}, \der{E} : G \Mapsto H$ be derivations with the same
  length, $\der{D}=\{\dder{D}_{i}\}_{i=0}^n$ and
  $\der{E}=\{\dder{E}_{i}\}_{i=0}^n$. If
  $\dder{D}_i \cdot \dder{D}_{i+1}$ is a switch of
  $\dder{E}_i \cdot \dder{E}_{i+1}$ for some $i \in [0,n-1]$ we write
  $\der{D} \leftrightsquigarrow_{(i,i+1)} \der{E}$. Then \emph{switch
    equivalence} is the least equivalence on derivation such that if
  $\der{D} \leftrightsquigarrow_{(i,i+1)} \der{E}$ then
  $\der{D} \equiv^s \der{E}$.
\end{definition}

\subsection{Relating sequential indepedence and switchability}

The fact that sequential indepedence implies switchability always
holds for linear rules. [REF] The result is so fundamental that it is
not even stated, in the sense that switchability is not introduced
axiomatically as in Definition~\ref{def:switch}, but it is based on
the explicit construction of a switch.

@@

\full{
  \begin{remark}
    \label{rem:fact}
    Notice that, since $f_0'$ is in $\mathcal{M}$, there is
    at most one $i_0' \colon L_0\to D_0'$ such that
    $m_0=f_0' \circ i_0'$. Moreover, we already know that
    there is at most one $i_1' \colon R_1\to D_1'$ such that
    $m_1'=f_1' \circ i_1'$, therefore, the
    independence pair $(i_1',i_0')$ between the component of a switch is
    uniquely determined.
  \end{remark}
}

@@

  Take now the set $S:=\{a,b,c\}$ and endow it with the smallest order
  such that $a\leq b$ and $a\leq c$.  Take also the set $\mathbb{N}$
  ordered by $\geq$, so that $0$ is its maximum.  Let $\X$ be the
  category associated to the order
  $(S, \leq)\oplus (\mathbb{N}, \geq)$ which, by \Cref{rem:iso} is
  $\mathsf{Iso}$-adhesive. Notice that in this category the arrows
  $(a,b)\colon a\to b$ $(a,c)\colon a\to c$ do not have a pushout
  because $b$ and $c$ do not have a supremum.  Define $\R$ as the set
  containing the following two rules
  \[\xymatrix{a & a \ar[r]^{(a, 0)} \ar[l]_{\id{a}} & 0 & a & a
      \ar[r]^{(a, b)} \ar[l]_{\id{a}} & b }\]
  We can then consider the following derivation
  $\der{D}=\{\dder{D}_i\}_{i=0}^1$ of length $2$, made by two
  sequential independent direct derivations.
  \[
    \xymatrix@C=30pt{a \ar[d]_{m_0}&& a \ar[d]_{k_0}\ar[ll]_{\id{a}}
      \ar[r]^{(a,0)} & 0 \ar@/^.35cm/[drrr]|(.315)\hole_(.45){\id{0}}
      \ar[dr]|(.3)\hole_{\id{0}} && a \ar@/_.35cm/[dlll]^(.45){(a,c)}
      \ar[dl]|(.3)\hole^{(a,0)}& a \ar[d]^{(a,0)}\ar@{>->}[l]_{\id{a}}
      \ar[rr]^{(a,b)} && b \ar[d]^{(b,0)}\\c &&
      \ar@{>->}[ll]^{(\id{c})} c \ar[rr]_{(c,0)}&& 0 &&
      \ar@{>->}[ll]^{\id{0}} 0 \ar[rr]_{\id{0}}&& 0}\]

  \noindent
  \parbox{10cm}{ \hspace{15pt}On the other hand, the rule
    applied by $\dder{D}_1$ cannot be applied to $c$. indeed the
    unique morphism $a\to c$ is $(a,c)$, thus we get a diagram
    aside. But $b$ and $c$ do not have a supremum in the poset
    underlying $\X$, thus we do not get a direct derivation from
    $c$.}  \parbox{3cm}{
    \vspace{-.3cm}$\xymatrix{a \ar[d]_{(a,c)}& a
      \ar[d]_{(a,c)}\ar@{>->}[l]_{\id{a}} \ar[r]^{(a,b)} & b \\c
      & c \ar[l]^{\id{c}} }$}  
\end{example}

Switches along the same independence pair are unique up to abstraction
equivalence (see Proposition~\ref{thm:switch_uni}).

\begin{example}
  \label{ex:seq-ind}
  Consider a rewriting system in $\cat{Graph}$, the category of
  directed graphs and graph morphisms, which is adhesive. The set of
  rules $\R = \{ \rho_0, \rho_1, \rho_2\}$ consists of three rules as
  depicted in Fig.~\ref{fi:rules} where numbers are used to represent
  the mappings from the interface to the left- and right-hand
  sides. Rules $\rho_0$ and $\rho_1$ are linear: $\rho_0$ generates a
  new node, while $\rho_1$ delete an edge. Instead, $\rho_2$ is non
  linear as it ``merge'' two nodes.

  \begin{figure}
    \input{Figure/rules.tex}

    \caption{Rules of the rewriting system grammar of the example.}
    \label{fi:rules}
  \end{figure}

  A derivation $\der{D}$ consisting of three steps $\der{D}_0$,
  $\der{D}_1$, $\der{D}_2$ each applying the corresponding rule
  $\rho_i$ is depicted in in Fig.~\ref{fi:derD}. The second and third
  steps $\der{D}_1$ and $\der{D}_2$ are clearly sequential
  independent, via an independence pairs which map nodes consistently
  with their numbering.\todo{P: having names for the graphs would
    probably help}

  \begin{figure}
    \input{Figure/derD.tex}
    \caption{The derivation $\der{D}$.}
    \label{fi:derD}
  \end{figure}

  The category $\cat{Graph}$ is tame by the result in ...  Thus,
  sequential independence ensure the existence of a switch $\der{E}$,
  depicted in Fig.~\ref{fi:derE}.
  \begin{figure}
    \input{Figure/derE.tex}
    \caption{The derivation $\der{E}$.}
    \label{fi:derE}
  \end{figure}
\end{example}
